---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Read zip file with read.table

```{r setoptions, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
activity <- read.table(unz("activity.zip", "activity.csv"),
                       header = TRUE,
                       sep = ",",
                       stringsAsFactors = FALSE,
                      )
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
1. Caculate the total number of steps per day
```{r}
activity_perday <- activity %>%
        group_by(date) %>%
        summarise(step_perday = sum(steps, na.rm = TRUE), .groups = "drop")
activity_perday
```

2. make a histogram of the total number of steps taken each day
```{r}
ggplot(data = activity_perday, aes(step_perday)) +
  geom_histogram() +
  labs(title = "Histogram of total number of steps taken ") +
  labs(x = "Steps per day")

```

3. Caculate mean and median of the total number of steps taken per day
```{r}
mean(activity_perday$step_perday, na.rm = TRUE)
median(activity_perday$step_perday, na.rm = TRUE)
```


## What is the average daily activity pattern?
1. use summarise to caculate the mean of each interval, and draw a time series plot.
```{r}
ac_pattern <- activity %>%
          group_by(interval) %>%
          summarise(step_average = mean(steps, na.rm = TRUE), .groups = "drop")
ggplot(data = ac_pattern, aes(interval, step_average))  +
  geom_line() +
    labs(title = "Daily activity pattern") +
    labs(x = "Intervals(5 minutes)", y = "Average steps")
```

2. find the interval has the max number of steps.
```{r}
which.max(ac_pattern$step_average)
``` 

## Imputing missing values
1. Caculate the total missing values in the dataset

```{r}
sum(is.na(activity$steps))
```
2. Impute strategy. There are days has all NAs, so I will use the mean for 5 minute interval to fill the data.
3. The new dataset name is ac_impute
```{r}
step_na <- is.na(activity$steps)
impute_ac <- function(activity_row) {
  if (is.na(activity_row[1])) {
     ac_pattern[ac_pattern[, 1] == as.integer(activity_row[3]), 2]
  } else{
     activity_row[1]
  }
}
ac_impute <- activity
ac_impute$steps <- as.numeric(unlist(apply(activity, 1, impute_ac)))

```
4. Make a histogram, and caculate the mean and median of the new dataset
```{r}
ac_impute_perday <- ac_impute %>%
        group_by(date) %>%
        summarise(step_perday = sum(steps, na.rm = TRUE), .groups = "drop")
ggplot(ac_impute_perday, aes(step_perday)) +
  geom_histogram() +
  labs(title = "Histogram of total number of steps taken (imputed dataset) ") +
  labs(x = "Steps per day")

```

mean and median of the new dataset.
```{r}
mean(ac_impute_perday$step_perday)
median(ac_impute_perday$step_perday)
```
According the two histogram, The zero number of total number of steps per day drops significantly, 
and the *mean* and *median* are both increased, which *mean* increased more. 

Because of the impute strategy. The intervals where  the steps was larger ,there will be more likely influenced by the impute,
The strategy of impute will  impact the pattern of the new dateset , the more data missing the more impact is.

## Are there differences in activity patterns between weekdays and weekends?

1. Creating new dataset as required
```{r}
ac_impute <- mutate(ac_impute, week = weekdays(date, abbreviate = TRUE))
ac_impute[ac_impute$week %in%
          c("Mon", "Tue", "Wed", "Thu", "Fri"), 4] <- "weekday"
ac_impute[ac_impute$week %in%
          c("Sat", "Sun"), 4] <- "weekend"
ac_impute$week <- as.factor(ac_impute$week)
```

2. make the plot.
```{r}
ac_impute_week <- ac_impute %>%
    group_by(interval, week) %>%
    summarise(step_interval = mean(steps, na.rm = TRUE), .groups = "drop")
ggplot(ac_impute_week, aes(interval, step_interval)) +
    facet_grid(week~.) +
    geom_line() +
    labs(title = "Patters between weekday and weekends") +
    labs(x = "Intervals(5 minutes)", y = "Average steps")


```
